---
title: "lab_01"
author: "Derek Willis"
date: "2023-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your class reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the UMD courses data.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
umd_courses <- read_rds("lab_01.rds")
```

## Answer questions

**Q1** Which department has the highest mean (average) number of seats for its classes? You need to show the department with the highest average first. Which one is the most surprising/interesting to you? **A1 It was interesting to me that engineering, cooperative education was the highest mean number of seats**

```{r}
umd_courses |>
  group_by(department) |>
  summarise(
    count_classes = n(),
    total_seats = sum(seats),
    mean_seats_a = mean(seats)) |>
    arrange(desc(mean_seats_a))
```

**Q2** Find out the greatest number of times any individual class has been offered in our dataset. Your code must show the greatest number of classes first. **A2**

```{r}
umd_courses |>
  group_by(id) |>
  summarise(
    count_classes = n()
  ) |>
  arrange(desc(count_classes))
```

**Q3** Count the number of courses for each instructor and department. Which instructor has taught the most? **A3 Keith Yearwood**

```{r}
umd_courses |>
  group_by(instructors, department) |>
  summarise( courses = n(),
 

  ) |>
  arrange(desc(courses)) |> na.omit()
```

**Q4** Given the results for Q3, what would have made it easier to find the answer? You don't have to know *how* to accomplish this; I want you to describe in words what would make the answer easier to see. How confident are you that the answer you got is the right one, and why? **A4:** I am moderately confident I got the right answer. Before, my code resulted in the instructors showing up in alphabetical order, but I did not want this so I included the desc() function in the arrange() function. Now my code looks like this, not in alphabetical order, but no way to be sure its in order of the instructor who taught the most. Using the desc() function makes me think I did it correctly but I would think the count_classes column would support my results too, which is why I am not 100% sure. It would be a lot easier if I knew whether the count_classes are supposed to show the amount of classes the instructors taught, and if so how do I make it show that.

**Q5** Take one of the questions you listed in the final question of the pre_lab and try to answer it here in code. If you don't know how to fully answer it, write what code you can and describe in words what else you would need to do. If you got an answer, is it interesting? **A5**

I used my third question, wondering the importance of the summarise () function. I wondered by couldn't we just group the columns and arrange them in the way we wanted to, why did we need to summarise? I tested it with just with term, department, and instructors. The entire chart popped up and when I used one specific column, the chart revolved around that column. For term it showed the only the term 202308, for department it got as far as departments beginning with W to U, and instructors was the same order as **Q3**. In all, there wasn't really a direction or order in the chart without the summarise() function. And when I put "term,department" within the group function, an error occured. In conclusion, we need the summarise () function to organize the codes and for the computer to properly recognize what we are asking of it.

```{r}
#
umd_courses |>
  group_by(term, department) |>
  arrange(desc( term, department))
```
