---
title: "lab_03"
author: "Derek Willis"
date: "2023-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)
# Load janitor
library(janitor)
```

## Load data and clean up the column names

Load the two datasets we worked with in the pre_lab: the Maryland payments to vendors and Maryland grants and loans and, using janitor, change the columns so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows

md_payments<- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2024.csv", guess_max=373564) |> clean_names()
  
  #read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2024.csv", guess_max=10)

  #read_csv("data/State_of_Maryland_Payments_Data_FY2008_to_FY2024.csv",guess_max=373564) |> clean_names()
  
# Load in the Maryland grants and loan data
md_grants_loans <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv") |> clean_names()

```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed. Write code to find rows where the vendor name contains the county's school system, starting with a broad search and then narrowing it down. When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

**A1:** 1204 observations of 8 variables.

```{r}
pg_schools <- md_payments |> 
  filter(str_detect(vendor_name,"PRINCE GEORGES PUBLIC") | 
    str_detect(vendor_name,"PRINCE GEORGES COUNTY SCHOOLS") | 
    str_detect (vendor_name,"PG COUNTY PUBLIC SCHOOLS") | 
    str_detect(vendor_name,"PRINCE GEORGES COUNTY SCHOOLS") | 
    str_detect (vendor_name,"	
PRINCE GEORGES") | 
    str_detect(vendor_name,"PRINCE GEORGES COUNTY PUBLIC SCHOOLS") |
    str_detect(vendor_name, "PRINCE G") |
    str_detect(vendor_name, "PRINCE GEORG") |
    str_detect(vendor_name, "PRINCE GEORGES COUNTY") |
    str_detect(vendor_name, "PRINCE GEORGE'S COUNTY") )
```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? What do you notice about the years listed, and what questions do you have about that data?

**A2:** 2022 with 1,523,657,781.06. What I noticed is the greatest amount changed depending on the category I used in the function. For example, with just 'fiscal_year' and vendor_name, it was 2023 with 78 count. And 'fiscal_year' and category was 2009 with 75 count. I wonder why it changed. I assume because the focus of the code changes but once I input amount, the count changes to majority 1 and 3.

```{r}
pg_schools|>
   group_by(`fiscal_year`, vendor_name, amount) |> 

  summarize(count = n()) |> 
  clean_names()|>
  arrange(desc(amount))
```

Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received. Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

A3. Cleaning ideas: Use case_when(), to remove doubles for specific things. For example, specifying the amount range you want or a specific number. I saw that fiscal_year 2008, and vendor_zip 20722 is duplicated by the amount is different. The vendor_name for that year and zip is also different which could be used in case_when() to remove a duplicate. Again for fiscal_year 2009 and vendor_zip 20722, but the category is the same but the amount is different - everything else is the same. Also, how the category is full of vendor payments, you would need to specify using the other columns to narrow down what you are looking for.

```{r}
pg_schools|>
  filter(str_detect(agency_name, "CAPITAL PROJECT"))|>
  arrange(vendor_name, vendor_zip,amount) |>

  clean_names()

```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which columns contain the information you want, your code should standardize the data needed to find just the grants for stem cell research in a single filter. Then create a total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

A4: I am not confident in my code because it looks it tells me more how much the grantor gave throughout that year, rather than the total of that year. I found it difficult to organize the amount in desc() function. But I can say certain grantors during a year were more generous than others.

```{r}
#md_grants_loans|> filter(str_detect(grantor, "Stem Cell Research Fund") | str_detect(grantor,"STEM CELL RESEARCH FUND"))|> group_by(amount, fiscal_year) |> summarise(amount) |> mutate(total = total (amount, "total")  ) |>

# cleaned_md_grants_loans <- md_grants_loans |> clean_names() |>  rename(total = grantor) |> 
   
 md_grants_loans |>
  clean_names() |> 
   group_by(grantor, "STEM CELL RESEARCH")|>
   summarise(amount, fiscal_year)
  rename(total = amount) |> 
  mutate(total = str_to_upper(total), grantee = str_to_upper(grantor))
 
arrange(desc(total))

#md_grants_loans |>  mutate(    total = sum(fiscal_year)/total(fiscal_year) |>  group_by(amount) |>  summarise(total_amount = sum(amount)) |>  arrange(desc(total_amount))
  

```
