---
title: "lab_04"
author: "Derek Willis"
date: "2023-01-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). Read the questions carefully and attempt to do all of what they ask. Make a list if it helps.

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load three packages for this: the tidyverse, lubridate and janitor.

**Task** load these three packages.

```{r}
# Turn off scientific notation
options(scipen=999)

library(tidyverse)
library(lubridate)
library(janitor)
```

For this lab, we want to investigate spending by Maryland state and local candidates during the 2022 election. For example, we could ask how much money have candidates and committees spent on advertising on Facebook? Which vendors received the most money for media expenses? We have the data, but it's messy - names of payees are spelled differently and capitalized differently - and we need to clean it up. We'll use a combination of RStudio and OpenRefine to do that.

The basic process here is to start in RStudio, export data so that we can use OpenRefine to clean it up, and then bring it back into RStudio to finish our analysis. The function to export data from RStudio is called `write_csv()` and it takes two arguments: the name of the dataframe and the path of the file to write it to, which is in quotes. Like `read_csv()`, but the variable name comes inside. Here's an example, in which you'd replace the dataframe and the path values.

```{r}
write_csv(dataframe, "folder/file.csv")
```

## Load Data

You'll need to load one data set, a CSV file of Maryland campaign expenditures located in the data folder called "maryland_expenses.csv"

**Task** Create a codeblock below, then read the data in and assign it to an appropriate variable name. You'll want to clean up the column names and make sure that any date columns are actually date datatypes.

```{r}
# Load the Maryland expenditures data table, clean the column names and convert columns with dates into actual dates.

md_expenses<- read_csv('data/maryland_expenses.csv') |> 
  clean_names()

md_expenses <- md_expenses |> mutate (expenditure_date = mdy(expenditure_date))
```

## Answer questions

**Q1.** You've been assigned to write about Maryland political committees' use of Facebook advertising. There's a category column in the data that has values for different kinds of spending. One of them is "Media". Write code to isolate Media expenses where the name of the payee is not NA. Then, using the `write_csv` function, make a CSV file of those media expenses in your data folder. Using OpenRefine, create a project using that CSV file, create a copy of the `payee_name` column called `payee_clean` and then standardize it, focusing your work on the payees that appear most often. Pay particular attention to variations of Facebook, bearing in mind that it changed its name in 2021: <https://about.fb.com/news/2021/10/facebook-company-is-now-meta/>).

Then, when you are done standardizing `payee_clean`, export it as a new CSV file to your data folder, *giving that file a different name* and read it back into RStudio as a new dataframe.

Then, using that `payee_clean` column, write code to find all of the spending where Facebook is the payee and calculate the total amount of spending for each committee and present the results showing the committee that spent the most first. Which committee is that?

**A1.** The top ten are: Moore Wes For Maryland, Bhandari Harry (H.B.) Friends of, Casa in Action PAC, Moore West For Maryland , King John for Governor, Franchot Anderson-Walker for Maryland Slate, Realtors PAC Md. (MD PAC), Casa in Action PAC, MSEA's Fund For Children And Public Education PAC, Blair David for Montgomery

```{r}
# One step at a time. Find the category column in the data that has values for different kinds of spending.

md_expenses |> 
  group_by(expense_category, committee_name) |>
  
  summarize(total = n())|>
  arrange(desc(total))

```

```{r}
# One of them is "Media". Write code to isolate Media expenses

media_expenses <- md_expenses |>
  filter(expense_category == 'Media') |>
  filter(!is.na(payee_name))


#where the name of the payee is not NA. They key here is using (!is.na(payee_name)) within a function


```

```{r}
# Export this table as a CSV

write_csv(media_expenses,"data/media_expenses.csv")


# Clean up work in OpenRefine
 media_expenses_clean <- read_csv('data/media-expenses_clean.csv.csv')



```

```{r}
# read the cleaned CSV 
# calculate top Facebook spenders

media_expenses_clean <- read_csv('data/media-expenses_clean.csv.csv')


```

**Q2.** Let's do some research on the top committees from question 1. Look up the top 10 spenders: if they are candidate committees, what offices were they running for? If they were not candidates, what issues do they seem to care about? Write some sentences answering those questions, then go the Facebook ad library ([https://www.facebook.com/ads/library/?active_status=all&ad_type=political_and_issue_ads&country=US&sort_data\\[direction\\]=desc&sort_data\\[mode\\]=relevancy_monthly_grouped&media_type=all](https://www.facebook.com/ads/library/?active_status=all&ad_type=political_and_issue_ads&country=US&sort_data%5Bdirection%5D=desc&sort_data%5Bmode%5D=relevancy_monthly_grouped&media_type=all){.uri}) and look for ads from the top spender. What do you notice about those ads?

**A2.** I looked up the top 10 spenders and everyone running for office sponsored themselves or were supported by organizations/PACs that supported them. And the organiztions/PACs advertised the politicians who supported in their cause. So the Realtors PAC md. advertised different politicians that supported pro-realtor efforts and the same for the MSEA and Casa for Action. MSEA stands for Maryland's Student Education Association, so they are pro student and education efforts in Maryland. Casa in action support progressive leaders. Interestingly though, Blair David's post have all be deleted.

**Q3.** Given the list of top spenders in A1 and the primary results for governor (<https://elections.maryland.gov/elections/2022/primary_results/gen_results_2022_1.html>), who is missing from that A1 list? Did those campaigns run Facebook ads, according to Facebook's ad library? Who are the biggest recipients of media expenses by the Democratic primary winner for governor (HINT: that person has two committees)?

**A3.** From the A1 list I see John King, Wes Moore, Franchot and Anderson Walker. All of these politicians did use Facebook ad library. Wes Moore is the one with the biggest recipients of media expenses.

```{r}

md_expenses |>
  group_by(payee_name, committee_name, amount)|>
  filter(payee_name == 'Facebook' ) |>
  summarize(payee_name)|>
  clean_names() |>
    
arrange(payee_name)
```

**Q4.** The media expense data you cleaned also has a standardized purpose column. Write code to generate a list of the committees with the total amount of spending on online advertising. How does this list compare to the one from Q1 about top Facebook spenders? Of the top 5 committees who spent the most on Facebook ads from Q1, are there any that appear to have spent *all* of their online advertising dollars on Facebook?

**A4.** Compared to the list from the Q1, I used the str_detect() function to specifically find results for variants of Facebook in the payee_name column. I am not sure if my function is able to find is the results from Q1 found all of their online advertising dollars on Facebook but I can say the one who spent the most was WFP National Pac - Non-Contribution Account at \$14000, which was not one of the results from Q1.

```{r}
md_expenses |>
  group_by(payee_name, committee_name, amount)|>
  filter(
    str_detect(payee_name, 'FACEBOOK')|
    str_detect(payee_name, 'Facebook') |
    str_detect(payee_name, 'Facebook Inc.')
  ) |>
  summarize(payee_name)|>
  clean_names() |>
    
arrange(desc(amount))
```

**Q5** You've been asked to summarize your findings about spending on Facebook by Maryland political candidates. How much do we know and what don't we know? What's the best story idea you've found here?

**A5:** One story idea I am thinking is how much is spent on social media ads and how effective they are, another could be how effective PACs are for running politicians (through how much is put into them, the tracked exposure on their ads, and whether they won or not). Before reviewing the links and charts I knew nothing about the ads, especially since I do not use Facebook. But this make me wonder with facebook owning instagram, could we see the same tracked ad library expanded there too?
