---
title: "lab_09.Rmd"
author: "derek willis"
date: "2023-04-12"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   A Datawrapper account
-   Data in CSV files in the `data` folder

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings
for this lab.

```{r}
# load libraries here
library(tidyverse) 
library(tidycensus)
library(ggthemes)
library(dplyr)
library(readr)
```

## Questions

**Q1.** You want to explore the differences between the electoral
performances of Dan Cox in 2022 and Larry Hogan in 2018. Except a map is
probably not the best way to show that. Let's make a chart instead, and
the type is up to you, but the goal is to clearly show the differences
between the two.

You MUST do the following things when answering this question: \* Load
the md_gov_county.csv and md_gov_county_18.csv files from the data
folder \* Join those dataframes, specifying the join column. \* Add
columns to generate percentages for Cox and Hogan. \* Export that joined
dataframe to a CSV file using `write_csv`. \* Using that exported CSV,
make a chart with Datawrapper. What kind of chart is up to you, so
experiment with a few kinds of charts to see what answers the question
above in the clearest way. Your chart MUST: \* Have a title, description
and source that tells readers about the data. The title and description
should highlight the most newsworthy or important aspect. \* Easily
readable labels and values \* Publish your chart \* Describe your
thought process below - why you chose to make the chart you did - and
add the URL of your chart.

**A1.** I did not run into any other problems aside from downloading the
joined data sets together as a csv for datawrapper to recognize it. I
chose the stacked bars graph because the main focus of this data is to
see the comparison between two politicians. It is easier to see them
directly next to each other than something else. My other ideas was a
line chart and pie charts. There you would also see the differences but
they were be separated. I prefer this way because they are right above
an another so you can really capture the difference in numbers.

link: <https://www.datawrapper.de/_/rK9Cq/>

```{r}
md_gov_county <- read_csv("~/JOUR472_fall2023/github_j472/lab_09/data/md_gov_county.csv")
View(md_gov_county)

md_gov_county_18 <- read_csv("~/JOUR472_fall2023/github_j472/lab_09/data/md_gov_county_18.csv")

View(md_gov_county_18)

```

```{r}

md_gov_county <- md_gov_county |>
  group_by(Cox)|>
  mutate(pct_cox = Cox/100)

md_gov_county_18 <- md_gov_county_18|>
  group_by(Hogan)|>
  mutate(pct_hogan = Hogan/100)

```

```{r}
md_gov_county_joined <- md_gov_county |> 
  left_join(md_gov_county_18, join_by("County"))

#md_gov_county_joined <- 'md_gov_county_joined'

write.csv(md_gov_county_joined, file = "md-gov-county-joined.csv")

view(md_gov_county_joined)


```

```{r}

(https://www.datawrapper.de/_/rK9Cq/ )
```

**Q2.** You're interested in seeing how the methods of voting have
changed in Maryland since 2018. Using the `md_votes_by_type.csv` that
has data from the top of the ticket (governor in 2018 and 2022,
president in 2020), make a chart showing the percentage of votes cast in
each category - early voting, election day and mail - over the three
years. You'll need to make those percentages in R and then export the
CSV using `write_csv`, then make the chart in Datawrapper. Have a title,
description and source that tells readers about the data. The title and
description should highlight the most newsworthy or important aspect.

HINT: In Datawrapper's "Check and Describe" tab, you can click on a
column to get the option of "hiding" it in any visualization.

What do you think the story here is? Put the answer below along with the
URL of your chart.

**A2.** The story is what new voting method has Maryland instated and
how is it now working, because the voting outcomes are decreasing as the
years progress.

link: <https://www.datawrapper.de/_/lmbEg/?v=2>

pct_votes_by_type

```{r}

md_votes_by_type <- read_csv("data/md_votes_by_type.csv")



pct_votes_by_type <- md_votes_by_type |>
  group_by(Total)|>
  mutate(pct_total = Total/100)

pct_votes_by_type
```

```{r}


#md_gov_county_joined <- 'md_gov_county_joined'

write.csv(pct_votes_by_type, file = "pct-votes-by-type.csv")

view(pct_votes_by_type)


```

**Q3.** You want to know more about the geography of 911 overdose calls
in Montgomery County, Maryland, in 2022. To do that, you'll make a
choropleth map in Datawrapper. You will need to:

-   Read the file into R and create columns for the number of calls in
    each zip code and the percentage of all calls that number represents
    and then export that CSV to use in your map.
-   Search for Maryland zip codes (use the 2020 data) as the base map.
-   Upload the CSV file.
-   Change the tooltip to show both the number and the percentage (you
    can put them side by side or put one under the other). Make sure to
    properly format both numbers to make them readable.
-   Add labels to your map that locate prominent places in the county.
-   Have a title, description and source that tells readers about the
    data. The title and description should highlight the most newsworthy
    or important aspect.
-   Make any other changes to data that helps you make a more accurate
    map.

What do you think the story here is, and what would you want to know
more about? Put the answer below along with the URL of your map.

**A3.** I would want to know about the demographic and socioeconomic
information of zipcode, 20886. The story definitely is that area having
more 911 overdose calls.

link: <https://www.datawrapper.de/_/yHRgX/>

```{r}

montgomery_2022_overdoses <- read_csv("data/montgomery_2022_overdoses.csv")

pct_overdose_calls <- montgomery_2022_overdoses|>
  group_by(zip, calltime_cleared)|>
  summarise(pct_calls = calltime_cleared / 100)
  
pct_overdose_calls
```

```{r}

#md_gov_county_joined <- 'md_gov_county_joined'

write.csv(pct_overdose_calls, file = "pct-overdose-calls.csv")

view(pct_overdose_calls)


```
