---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

marvel_data <- read_csv("Marvel Group Project/marvel-wikia-data.csv")


dc_data <- read_csv("Marvel Group Project/dc-wikia-data.csv")



# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

#dc_data <- dc_data |> mutate(FIRST APPEARANCE = strftime(`FIRST APPEARANCE`, mdy(date))) |>show_col_types = FALSE



  

```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**:

    Women were underrepresented among characters with at least one appearance, and females only made up 29.3% of the DC character list and 24.7% of the Marvel roster.

-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

#dc_data <- dc_data |> group_by(SEX, APPEARANCES) |> mutate(pct_representation = Female Characters/`FIRST APPEARANCES`)*100

#dc_data <- dc_data|> mutate(percent_female = (100*n()/NROW ("Female Characters")))|> arrange (percent_female)
 

dc_data |>
  group_by(SEX)|>
  summarize(count = n())|>
  mutate(pct_female = count/6896*100)|>
  arrange(desc(pct_female))
  
marvel_data |>
  group_by(SEX)|>
  summarize(count = n())|>
  mutate(pct_female = count/16376*100)|>
    arrange(desc(pct_female))
  

# Display results of code below this codeblock



```

### Sentence 2

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
